---
title: "rna_seq_project"
author: "Jose Chacon (PID A59010515)"
date: "11/19/2021"
output: pdf_document
---

```{r message=FALSE}
library(DESeq2)
```

# Data import
```{r}
metaFile <- "GSE37704_metadata.csv"
countFile <- "GSE37704_featurecounts.csv"

# Import metadata and take a peak
colData = read.csv(metaFile, row.names=1)
head(colData)

```

```{r}
# Import countdata
countData = read.csv(countFile, row.names=1)
head(countData)
```

```{r}
# Note we need to remove the odd first $length col
countData <- as.matrix(countData[,2:7])
head(countData)
```




#Barry code for removing code
#counts <- as.matrix(countData[, -1])

#Remove zeros (use previous code)
#inds <- which(meancounts[,1:2] == 0, arr.ind=TRUE)
#head(inds)
#to.rm <- unique(sort(inds[,"row"]))
#mycounts <- meancounts[-to.rm,]
#head(meancounts[-to.rm,])
```{r}
inds <- which(countData > 0, arr.ind=TRUE)
to.rm <- unique(sort(inds[,"row"]))
mycounts <- countData[to.rm,]
head(countData[to.rm,])
```
***CODE
```{r}
df2 <- countData[rowSums(countData[]) > 0, ]
nrow(df2)
```

```{r}
#new_countData = countData(rowSums([, -1])>0) )
#head(new_countData)
```



```{r}
nrow(mycounts)
```
```{r}
counts <- read.csv("GSE37704_featurecounts.csv", row.names=1)
```

```{r}
countData=read.csv(countFile, row.names=1)
```

```{r}
countData <- as.matrix(countData[,-1])
```

```{r}
counts <- countData[rowSums(counts) !=0,]
```

#Barry code 
```{r}
pca <- prcomp(t(counts))
summary(pca)
```

#PCA
```{r}

```

```{r}
#pca <- prcomp(mycounts , scale=TRUE) old code
```

```{r}
plot(pca$x[,1:2])
```

```{r}
pca <- prcomp(t(counts))
plot(pca$x[,1:2], pch=16, col=as.factor(colData$condition))
text(pca$x[,1:2], labels = colData$condition)
```
```{r}
library(ggplot2)

x <- as.data.frame(pca$x)
x$condition <- colData$condition

ggplot(x) +
  aes(PC1, PC2, col=condition) +
  geom_point()
```

```{r}
dds = DESeqDataSetFromMatrix(countData=counts, #change to nonzero data
                             colData=colData,
                             design=~condition)
dds = DESeq(dds)
```

```{r}
dds <- DESeq(dds)
res <- results(dds)
res
```

#check similarity later
```{r}
#mycounts == df2
```

```{r}
summary(res)
```

```{r}
plot( res$log2FoldChange, -log(res$padj) )

```
```{r}
# Make a color vector for all genes
mycols <- rep("gray", nrow(res) )

# Color red the genes with absolute fold change above 2
mycols[ abs(res$log2FoldChange) > 2 ] <- "red"

# Color blue those with adjusted p-value less than 0.01
#  and absolute fold change more than 2
inds <- (res$listDATA$pvalue <0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"

plot( res$log2FoldChange, -log(res$padj), col=mycols, xlab="Log2(FoldChange)", ylab="-Log(P-value)" )
```
```{r}
#mycols <- rep("gray", nrow(res))
#mycols[abs(res$log2FoldChange)>2] <- red
#mycols[res$pad > 0.05] <- "gray"

#plot(res$log2FoldChange, -log(res$padj), col=mycols)

```

```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")

columns(org.Hs.eg.db)

res$symbol = mapIds(org.Hs.eg.db,
                    keys=row.names(res), 
                    keytype="ENSEMBL",
                    column="SYMBOL",
                    multiVals="first")

res$entrez = mapIds(org.Hs.eg.db,
                    keys=row.names(res),
                    keytype="ENSEMBL",
                    column="ENTREZID",
                    multiVals="first")

res$name =   mapIds(org.Hs.eg.db,
                    keys=row.names(res),
                    keytype="ENSEMBL",
                    column="GENENAME",
                    multiVals="first")

head(res, 10)
```

#Pathway
```{r}
library(pathview)
```

```{r}
library(gage)
library(gageData)

data(kegg.sets.hs)
data(sigmet.idx.hs)

# Focus on signaling and metabolic pathways only
kegg.sets.hs = kegg.sets.hs[sigmet.idx.hs]

# Examine the first 3 pathways
head(kegg.sets.hs, 3)
```

```{r}
foldchanges = res$log2FoldChange
names(foldchanges) = res$entrez
head(foldchanges)
```

```{r}
# Get the results
keggres = gage(foldchanges, gsets=kegg.sets.hs)
```

```{r}
attributes(keggres)
```

```{r}
# Look at the first few down (less) pathways
head(keggres$less)
```

```{r}
pathview(gene.data=foldchanges, pathway.id="hsa04110")

```

```{r}
# A different PDF based output of the same data
pathview(gene.data=foldchanges, pathway.id="hsa04110", kegg.native=FALSE)
```

```{r}
## Focus on top 5 upregulated pathways here for demo purposes only
keggrespathways <- rownames(keggres$greater)[1:5]

# Extract the 8 character long IDs part of each string
keggresids = substr(keggrespathways, start=1, stop=8)
keggresids
```

```{r}
pathview(gene.data=foldchanges, pathway.id=keggresids, species="hsa")

```

```{r}
data(go.sets.hs)
data(go.subs.hs)

# Focus on Biological Process subset of GO
gobpsets = go.sets.hs[go.subs.hs$BP]

gobpres = gage(foldchanges, gsets=gobpsets, same.dir=TRUE)

lapply(gobpres, head)
```

```{r}
sig_genes <- res[res$padj <= 0.05 & !is.na(res$padj), "symbol"]
print(paste("Total number of significant genes:", length(sig_genes)))
```

```{r}
write.table(sig_genes, file="significant_genes.txt", row.names=FALSE, col.names=FALSE, quote=FALSE)
```

```{r}
sessionInfo()
```

