---
title: "class_17"
author: "Jose Chacon (PID A59010515)"
date: "11/24/2021"
output:
  pdf_document:
    latex_engine: xelatex
---

## Background

In this begore thanksgiving class when many of our classmates are traveling
let's have a look at COVID-19 vaccination rates around the State.

Vaccination rates from 
https://data.ca.gov/dataset/covid-19-vaccine-progress-dashboard-data-by-zip-code

## Import data
```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```

>Q. How many entries do we have?

```{r}
nrow(vax)
```

We can use the **skimr** package and the `skim()` function to get a quick overview of the structure of this dataset

```{r}
skimr::skim(vax)
```

Notice that one of these columns is a date column, Working with time and dates get's annoying quickly. We can use **lubridate** package to amek this easier

#install.packages("tidyverse") run in cosole
```{r}
library(lubridate)
today()
```

> Q. How many days since the first entry in this dataset?

```{r}
vax$as_of_date[1]
```
Tnis will not work vecause our data column was read as character..
```{r}
#today() - vax$as_of_date[1]
```

```{r}
d <- ymd(vax$as_of_date)
```

```{r}
today() - d[1]
```

I will make the *as_of_date* colum Date format

```{r}
vax$as_of_date <-  ymd(vax$as_of_date)
```


> Q. When was the dataset last updated? What is the last date in this dataset? How many days since the last update?

```{r}
today() - vax$as_of_date[nrow(vax)]
```

>Q How many days does the dataset span?

```{r}
vax$as_of_date[ nrow(vax)] - vax$as_of_date[1]
```

>Q. How many different ZIP code areas are there

```{r}
example <-  vax[2]
head(example)
```
```{r}
uni_example <-  unique(example)
#uni_example

nrow(uni_example)
```
BArry codde

```{r}
length(unique(vax$zip_code_tabulation_area))
```


To work with ZIP codes we can use the **zipcodeR** 
#install install.packages("zipcodeR", dependency= T)
```{r}
library(zipcodeR)

reverse_zipcode(c('92037', "92109") )
```

## Focus in on San Diego County

We want a subset the full CA `vax` data down to just San Diego County

We could do this with base R
```{r}
inds <- vax$county == "San Diego"
nrow(vax[inds,])
```

Subsetting can get tedious and complicated quickly when you have multiple things we want to subset by.

```{r}
library(dplyr)
```

We will use the `filter()` function to do our subsetting from now on.


We want to focus in on San Diego County

```{r}
sd <- filter(vax, county=="San Diego")
nrow(sd)
```

```{r}
sd.20 <- filter(vax, county=="San Diego",
       age5_plus_population> 20000)

nrow(sd.20)
```

How many unique zipcodes in San diego county

```{r}
length(unique(sd))
```


> Q. WHat is the average vaccination rate of San Diego county as of yesterday?

```{r}
sd.now <- filter(vax, county=="San Diego",
                 as_of_date=="2021-11-23")
head(sd.now)
```



```{r}
sd.now$percent_of_population_fully_vaccinated
```

```{r}
summary(sd.now$percent_of_population_fully_vaccinated)
```

>Q. Make a histogramof these values

```{r}
hist(sd.now$percent_of_population_fully_vaccinated)
```



This plot above is going to be susciptible to being skewed by the ZIP code areas with small populations. These will have a big effect for just a small number of unvax-ed folks....

>Q. What is the population iof the 92037 ZIP code area?

```{r}
lj <- filter(sd.now, zip_code_tabulation_area=="92037")
lj$age5_plus_population
```

>Q. WHat is the average vaccination value for this UCSD/La Jolla code area?

```{r}
lj$percent_of_population_fully_vaccinated
```

>Q. What about the ZIP code 92122


```{r}
lj2 <- filter(sd.now, zip_code_tabulation_area=="92122")
lj2$age5_plus_population
```

```{r}
lj2$percent_of_population_fully_vaccinated
```

>Q. 92124

```{r}
filter(sd.now, zip_code_tabulation_area=="92124")
```

TIme series of vaccination rate for a given ZIP code area.
Start with 92037.
```{r}
lj <- filter(vax, zip_code_tabulation_area=="92037")

library(ggplot2)
```

```{r}
ggplot(lj) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1)
  labs(x="Date", y="Percent Vaccinated")
```

Let's make this plot for all Sand Diego County ZIP code areas that have a population as least as large as 92037/

```{r}
sd.36 <- filter(vax, county=="San Diego", 
             age5_plus_population> 36144)
head(sd.36)
```


>Q. How many ZIP codes areas in San Diego county have a population larger than 92037?

```{r}
length(unique(sd.36$zip_code_tabulation_area))
```

Let's make a plot

```{r}
ggplot(sd.36) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated,
      group=zip_code_tabulation_area ) +
  geom_line()
  labs(x="Date", y="Percent Vaccinated")
```


>Q. Make a plot like this for all ZIP code areas in the stete with at least as large as La Jolla.

```{r}
ca <- filter(vax, age5_plus_population > 36144)
```

>Q. How many 

```{r}
length(unique(ca$zip_code_tabulation_area))
```
Make our plot

```{r}
ggplot(ca) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated,
      group=zip_code_tabulation_area ) +
  geom_line(alpha=0.2) +
  geom_hline(yintercept = 0.67, color="red")+
  labs(x="Date", y="Percent Vaccinated")

```

>Q. What is the mean across the state for the 36k+ population code?

```{r}
ca.now <- filter(ca, as_of_date=="2021-11-23")
summary(ca.now$percent_of_population_fully_vaccinated)
```

```{r}
home <- filter(vax, county=="Los Angeles", zip_code_tabulation_area=="90029")
home.now <-  filter(home, as_of_date =="2021-11-23")
summary(home.now$percent_of_population_fully_vaccinated)
```


---













