---
title: 'class 6: R function'
author: "Jose Chacon (PID A59010515)"
date: "10/15/2021"
output:
  pdf_document: default
  html_document: default
---

## Quick Rmarkdown intro

We can write text of course just like any file. We can **style text to be bold** or *italic*:

Do:

-this
-and that
-and another thing

THis is more text
and this is a new line

---

We can include some code:

```{r}
plot(1:10)
```

## Time to write a function

>**Q1**. Write a function grade() to determine an overall grade from a vector of student homework
assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an
NA value) this can be used as a score to be potentially dropped. Your final function should be
adquately explained with code comments and be able to work on an example class gradebook
such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]

```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```

Find lowest score, to exclude from average. Lowest score can be found with min() and position can be found with which.min()
```{r}
which.min(student1)
```

# add '-' in front to exclude data in vector, in this case 'which.min' , excluding the lowest score from the vector.

```{r}
student1[ -which.min(student1) ]
```

Now **mean()** can function can be used to get average

```{r}
mean(student1[-which.min(student1) ])
```

Does it work for student2?

```{r}
mean(student2[-which.min(student2) ])
```
```{r}
which.min(student2)
```
```{r}
mean(student2)
```

```{r}
mean(student2, na.rm=TRUE)
```
```{r}
mean(student3, )
```

Replace NA value with 0


```{r}
student3[is.na(student3)] <- 0
print(student3)
```

na() makes it a binary T or F at each position

```{r}
is.na(student2)
```

! turns it upside down, reverses it
```{r}
!is.na(student2)

```

replace NA values with zero and conserve original
```{r}
student.prime <- student2
student.prime[is.na(student.prime)] = 0
student.prime
```

get mean excluding lowest score

```{r}
mean(student.prime[ -which.min(student.prime)])
```
 
 student 3?
 
```{r}
student.prime <- student3
student.prime[is.na(student.prime)] = 0
student.prime
mean(student.prime[ -which.min(student.prime)])


```

Clear up code
```{r}
x <- student3
x[is.na(x)] = 0
x
mean(x[ -which.min(x)])
```
 
 
 New student
```{r}
student4 <- c(100, NA, 90, "90", 90, 90, 97, 80)
student4
new_student4<- as.numeric(student4)
new_student4
```
 
```{r}
x <- new_student4
x[is.na(x)] = 0
x
mean(x[ -which.min(x)])
```
Write function 
function are composed of 3 things
a name, input arg, and a body
```{r}
grade <- function(x) {
  x <- as.numeric(x)
  x[ is.na(x)] = 0
  mean(x [ -which.min(x)])
}
```
 
```{r}
grade(student2)
```

Now grade the class

```{r}
gradebook <- "https://tinyurl.com/gradeinput"
scores <- read.csv(gradebook, row.names=1)
scores
```

Use the  **apply()**  fuction to grade students with the **grade()** function
>**Q1**

```{r}
ans <- apply(scores, 1, grade)
ans
```

> **Q2**  who is top scoring student

```{r}
which.max(ans)
```
## student 18 scored the highest
Q3. which test was toughest

```{r}
q3 <- apply(scores, 2, mean)
q3
```

ignore NA values # not great since it means if someone didn't turn it it would be zero
```{r}
apply(scores, 2, mean, na.rm=TRUE)
```
replace/mask NA values to zero
```{r}
mask <- scores
is.na(mask)
mask[is.na(mask)] <- 0
mask
```


Now use apply utilizing the "masked" scores!
```{r}
apply(mask, 2, mean)

```
## Hardest test was hw2
Q4.Q4. Optional Extension: From your analysis of the gradebook, which homework was most
predictive of overall score (i.e. highest correlation with average grade score)? [1pt]

Here use the **cor()** function
```{r}
cor(mask$hw5, ans)
```

```{r}
cor(mask, ans)
```

can call **cor()** for every hw and get a value, use apply to do them all

```{r}
ans
apply(mask, 2, cor, ans) 
```

## hw5 is most predictive


## NEW
```{r}
grade2 <-function(x, drop.lowest=TRUE){
  x <- as.numeric(x)
  x [ is.na(x)] = 0
  if(drop.lowest) {
    mean(x[-which.min(x)])
  } else {
    mean(x)
  }
} 
```

```{r}
student <- c(100, 100, 100, 90)
grade2(student, drop.lowest = FALSE)
```

